Running `Earth_obs_cleaning` on MCD19A2: Results from CONUS

### 2022-04-26 R/CONUS_AOD.Rmd
cleanup mapping to noaa regions; closes #1

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r basics, echo=FALSE}
library(targets)
library(data.table)
library(DT)
library(patchwork)
library(ggplot2)
library(sf)
library(knitr)
library(kableExtra)
# library(flextable)
# set_flextable_defaults(fonts_ignore=TRUE)
```

@@

# Contiguous US state boundaries
# will replace with a region defined by Kodi later
usa = st_as_sf(maps::map("state", fill = TRUE, plot = FALSE))
```

```{r get_noaa_regions, include=FALSE}
# assign AERONET stations to states
tar_load(aer_stations)
aer_stations_sf <- st_as_sf(aer_stations, coords = c("lon", "lat"), crs = "epsg:4326")
sf::sf_use_s2(FALSE) # we were getting errors about the USA geometry otherwise
aer_stations_sf <- st_join(aer_stations_sf, usa)
dt <- merge(dt, aer_stations_sf, by = "Site_Name")
# some Site_Names don't fall within land polygons for whatever reason
dt[is.na(ID), table(Site_Name)]
# where are the sites that didn't fall within a state?
site_nostate <- st_as_sf(dt[is.na(ID), ][unique(Site_Name), .(Site_Name, `Site_Elevation(m)`, geometry), mult = "first"], crs = "epsg:4326")
# ggplot() +
#   geom_sf(data = usa, fill = NA) +
#   geom_sf(data = site_nostate, color = "red")
# what proportion of data are not in a state?
dt[, proportions(table(is.na(ID)))]
# assign these to the closest state
matched <- st_join(site_nostate,
                   usa,
                   join=nngeo::st_nn, k=1)
# plot these matchups
# ggplot() +
#   geom_sf(data = usa, fill = NA) +
#   geom_sf_text(data = matched, aes(label = toupper(substr(ID, 1, 1))), color = "red") + theme_minimal()
dt[as.data.table(matched, key = "Site_Name"), ID := i.ID]

# where do our data come from (by State)?
dt[, .(uniqueN(Site_Name), .N), by = ID][order(N, decreasing = T)]

# trying to get from state names to state abbreviations
# ugly capitalization
dt[ , state.abb := state.abb.plusdc[match(stringr::str_to_title(ID), c(state.name, "District Of Columbia"))]]
dt[, state.noaa.region := factor(
   sapply(state.abb, USE.NAMES = F,
       function(sa) if (sa %in% c("AK", "HI")) NA else
           names(noaa.regions)[which(sapply(noaa.regions, function(r)
               sa %in% r))]),
   levels = names(noaa.regions))]
```

@@

# map by NOAA regions

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = here::here())
```

```{r basics, echo=FALSE}
suppressPackageStartupMessages(
  {library(targets)
    library(Just.universal)
    library(data.table)
    library(DT)
    library(patchwork)
    library(ggplot2)
    library(sf)
    library(Just.universal)
    library(knitr)
    library(kableExtra)
    # library(flextable)
  })
# set_flextable_defaults(fonts_ignore=TRUE)
```

@@

# Contiguous US state boundaries
# will replace with a region defined by Kodi later
conus = get_conus()
```

```{r get_noaa_regions, include=FALSE}
# assign AERONET stations to states
tar_load(aer_stations)
aer_stations_sf <- st_transform(crs = st_crs(conus), st_as_sf(aer_stations, coords = c("lon", "lat"), crs = "epsg:4326"))
aer_stations_sf <- st_join(aer_stations_sf, conus)
dt <- merge(dt, aer_stations_sf[, c("Site_Name", "STUSPS")], by = "Site_Name")
# some Site_Names don't fall within land polygons for whatever reason
dt[is.na(STUSPS), table(Site_Name)]
# where are the sites that didn't fall within a state?
site_nostate <- st_as_sf(dt[is.na(STUSPS), ][unique(Site_Name), .(Site_Name, `Site_Elevation(m)`, geometry), mult = "first"])
# ggplot() +
#   geom_sf(data = conus, fill = NA) +
#   geom_sf(data = site_nostate, color = "red")
# what proportion of data are not in a state?
dt[, proportions(table(is.na(STUSPS)))]
# assign these to the closest state
matched <- st_join(site_nostate,
                   conus,
                   join=nngeo::st_nn, k=1)
# plot these matchups
# ggplot() +
#   geom_sf(data = conus, fill = NA) +
#   geom_sf_text(data = matched, aes(label = STUSPS), color = "red") + theme_minimal()
dt[as.data.table(matched, key = "Site_Name"), STUSPS := i.STUSPS]

# where do our data come from (by State)?
dt[, .(uniqueN(Site_Name), .N), by = STUSPS][order(N, decreasing = T)]

# assign NOAA region from state USPS codes
dt[, state.noaa.region := factor(
   sapply(STUSPS, USE.NAMES = F,
       function(sa) if (sa %in% c("AK", "HI")) NA else
           names(noaa.regions)[which(sapply(noaa.regions, function(r)
               sa %in% r))]),
   levels = names(noaa.regions))]
```

@@

## map by NOAA regions
# map_regions <- ggplot(conus) + geom_sf()
# print(map_regions)
```
